---
title: "DCM_family_EHJ-HF"
author: "Christoffer Vissing"
date: "26/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(networkD3)
library(webshot)
```

## Family screening in DCM

This document contains the data and figures for the manuscript ....

### Loading data and creating variables
```{r Data load}
b <- read_excel("H:/01 PhD/02 Precision diagnostics/Projekt 05 - Family screening yield/DCM_fam_yield_v1.1.xlsx", 
                sheet = "Collected", na = "NA")
f <- read_excel("H:/01 PhD/02 Precision diagnostics/Projekt 05 - Family screening yield/DCM_fam_yield_v1.1.xlsx", 
                sheet = "Family", na = "NA")
```
In the following section/chunk variables are created or altered
```{r Variable wrangling}

b <- b %>% dplyr::mutate(BMI = Kg1/(Height/100)^2,
                         NYHA_1st = factor(NYHA_1st),
                #ECHOVARIABLES at baseline (DILATION LVEF SO ON)
                         dil0 = 100*(LVED0/LVED0_pred)-100, #different from predicted value
                         Dil_0 = ifelse(dil0>12, 1,0), # LV dimensions larger than 112% predicted
                         Dil_0_17 = ifelse(dil0>17, 1,0), # LV dimensions larger than 117% predicted
                         LVEF0_reduced = ifelse(LVEF0<60, 1,0), #Variable if LVEF less than 60
                #ECHOVARIABLES at end vs diagnosis
                         delta_LVEF = LVEF2-LVEF1, #Change in LVEF from baseline/diagnosis to end of follow-up
                         delta_LVEDi = LVEDi2-LVEDi1, #Change in indexed LVED from baseline/diagnosis to end of follow-up
                #GENETIC VARIABLES
                         Known_genofam = if_else(Geno_fam=="Y", "Yes", "No", missing = "Unknown"),
                         genopositive =  if_else(Geno_pos=="Y", "Yes", "No", missing = "Unknown"),
                         genopositive_p_lp =  if_else(Geno_pos_p_lp=="Y", "Yes", "No", missing = "Unknown"),
                         genes = if_else(Gene=="TTN", "TTN", 
                                  if_else(Gene=="RBM20", "RBM20", 
                                          if_else(Gene=="LMNA", "LMNA",
                                                  if_else(Gene=="MYH7", "MYH7",
                                                          if_else(Gene=="DSP", "DSP", 
                                                                  if_else(Geno_fam=="Y","Other",
                                                                          if_else(Geno_fam=="N","Negative", "Unknown"))))))),
                 #ECG
                         QRS_over_120 = if_else(QRS1>120, 1, 0),# QRS >120
                         QRS_over_110 = if_else(QRS1>110, 1, 0),# QRS >110
                         Tinv = if_else(Tinv_Ant1+Tinv_Inf1+Tinv_Lat1>0.5, 1, 0), #T-wave inversion
                         longQTc = if_else(Sex=="M", if_else(QTc1>470,1,0), if_else(QTc1>480,1,0)),
                         QTc_450 = if_else(QTc1>450,1,0),
                         Abnorm_ECGn = if_else(Tinv==1,1,0, missing = 0)+
                           if_else(QRS_over_120==1,1,0,missing = 0)+
                           if_else(PR1>200,1,0, missing = 0)+
                           if_else(`2AVB1`==1,1,0, missing = 0)+
                           if_else(`3AVB1`==1,1,0, missing = 0)+
                           if_else(AFLI1==1,1,0, missing = 0)+
                           if_else(longQTc==1,1,0, missing = 0),
                         Abnorm_ECG = if_else(Abnorm_ECGn>0.5, 1,0),
                         Abnorm_TTEn = if_else(LVEF0_reduced==1,1,0, missing = 0)+
                           if_else(Dil_0_17==1,1,0, missing = 0),
                         Abnorm_TTE = if_else(Abnorm_TTEn>0.5, 1,0),
                         Abnorm_TTE_or_ECG = if_else(Abnorm_ECG+Abnorm_TTE>0.5,1,0),
                         Abnorm_TTE_and_ECG = if_else(Abnorm_ECG+Abnorm_TTE>1.5,1,0)
                )

d<- b %>% filter(Proband!="Y")

# Only relatives w/o FDC at baseline
d3 <- filter(d, CM_first_visit==0)

# Only relatives w/o FDC at baseline and who are not geno-type-neg
d3_XGneg <- filter(d3, genopositive_p_lp!="No")

#Only relatives without acute FDC at baseline
d4 <- d %>%   mutate(dummy = if_else(d$Ambulatory=="Acute",1,0,missing = 0),
         dummy = dummy+ CM_first_visit,
         dummy = if_else(dummy>1.5, 0,1,missing = 1)) %>% 
  filter(dummy==1)

#MAKING GENETABLE
f1 <- filter(f, Gene!="NA")
f1<- filter(f1, Gene!="None")
f1.table <- dplyr::select(f1, Gene, Gene_variant, ACMG_call, ACMG_criteria)
f1.table <- arrange(f1.table, Gene)
f1.table %>% kableExtra::kbl()%>%
  kableExtra::kable_paper(full_width = TRUE) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "condensed", "responsive")) %>%
  kableExtra::scroll_box(width = "800px", height = "3600px") #%>%
  #kableExtra::save_kable("VUFFI.png")
```

The following figure is a Sankey diagram showing the flow (outcomes) of patients according to baseline findings
```{r}
########################################
# Figure 3
##########################################
graphd <- 
  d  %>%  
#Exclude subjects diagnosed with FDC at baseline
  filter(CM_first_visit!=1) %>% 
#Select proper variables
  select(CM_c,Abnorm_TTE_or_ECG, genopositive_p_lp) %>% 
#Create relevant variables for plotting/data wrangling
  mutate(genopositive = case_when(genopositive_p_lp=="Yes"~1,
                                  genopositive_p_lp=="No"~0,
                                  TRUE~NA_real_),
         group = factor(case_when(genopositive==1~"G+",
                           genopositive==0~"G-",
                           Abnorm_TTE_or_ECG==1~"Abnormal",
                           Abnorm_TTE_or_ECG==0~"Normal",
                           TRUE~NA_character_), levels = c("Normal", "G-", "G+", "Abnormal"))) %>%  
  group_by(group) %>% 
#Summarise findings
  summarise( FDC= sum(CM_c), Healthy = sum(CM_c==0)) %>% 
  pivot_longer(cols = c(FDC, Healthy)) %>% 
  mutate(name = factor(name, levels = c("Healthy", "FDC")),
         source = case_when(group=="G-"~0,
                            group=="Normal"~1,
                            group=="Abnormal"~2,
                            group=="G+"~3),
         target = case_when(name=="FDC"~5,
                            name=="Healthy"~4),
         group_color = factor(case_when(name=="FDC"~"l1",
                                        name=="Healthy"~"l2"))) %>% 
#To avoid "empty" links
  filter(value>0)
  
#Create dataset for nodes
grnodes <- data.frame( node = c(0,1,2,3,4,5), 
                       name = c("","","","","","")) %>% 
  mutate(group = factor(c("G-","Normal","Abnormal","G+","Healthy", "FDC")))
#----------------
#Choose colors for plot
my_color <-  'd3.scaleOrdinal() .domain(["Normal","G-","G+","Abnormal","Healthy", "FDC","l1", "l2"]) .range(["#001959", "#184E60", "#577646", "#B28C32", "#9fbd8f", "#fca1a1", "#fdcece", "#d9d9d9"])'
#--------------
#Relevant size plot
Fig3 <- sankeyNetwork(Links = graphd, Nodes = grnodes, Source = "source",
                      Target = "target", Value = "value", NodeID = "name",
                      units = "Relatives" , fontSize = 14, nodeWidth = 80, 
                      colourScale = my_color, NodeGroup = "group", LinkGroup = "group_color",
                      nodePadding = 30, fontFamily = "sans",
                      height = 600, width = 400, iterations = 400, sinksRight = T)

Fig3
#Save as HTML
saveNetwork(Fig3, "Fig3.html")
library(webshot)
#Save as png
webshot("Fig3.html", "VUF.png")

#Bigger Sankey for improved resolution
Fig3 <- sankeyNetwork(Links = grdf, Nodes = grnodes, Source = "source",
                      Target = "target", Value = "value", NodeID = "name",
                      units = "Relatives" , fontSize = 140, nodeWidth = 800, 
                      colourScale = my_color, NodeGroup = "group", LinkGroup = "group",
                      nodePadding = 300, fontFamily = "sans",
                      height = 6000, width = 4000, iterations = 400, sinksRight = TRUE)
saveNetwork(Fig3, "CI.html")
webshot("CI.html", "EHJ-HF Figure 3.png")
```


**Prevalence in male and female relatives**
```{r}
###############################################---------------------------------#
library(epiR);library(survival); library(tidyverse);library(ggtext)
#----------------------------------#
# Figure 2 panel A
#---------------------------------#
fig2a <- 
  b %>%
#Select relatives and exclude those diagnosed prior to date of first screening
  filter(Proband != "Y" &
           1.7 > (CM_first_visit + if_else(Ambulatory == "Acute", 1, 0, missing = 0))) %>%
#Creating age-groups
  mutate(Agegroup = cut(
    age_baseline,
    right = TRUE,
    c(-Inf, 30, 45, 60, Inf),
    labels = c("<30", "30-44", "45-59", "\u226560")
  )) %>%
  group_by(Sex, Agegroup) %>%
  summarise(npop = n(), ncas = sum(CM_first_visit)) %>%
  ungroup()  %>%
  select(ncas, npop) %>%
  as.matrix() %>%
#Computing prevalence and exact confidence intervals
  epi.conf(ctype = "prevalence", method = "exact") %>% as_tibble() %>%
#Fuse with original data
  bind_cols(
    b %>% filter(Proband != "Y" &
                   1.7 > (
                     CM_first_visit + if_else(Ambulatory == "Acute", 1, 0, missing = 0)
                   )) %>%
      mutate(Agegroup = cut(
        age_baseline,
        right = TRUE,
        c(-Inf, 30, 45, 60, Inf),
        labels = c("<30", "30-44", "45-59", "\u226560")
      )) %>%
      group_by(Sex, Agegroup) %>%
      summarise(npop = n(), ncas = sum(CM_first_visit)) %>%
      ungroup()
  )  %>%
  mutate(sur = 1 - est,
         Sex = fct_recode(Sex, women = "F", men = "M")) %>%
#PLOTTING
  ggplot(aes(x=Agegroup, y=est*100, group = Sex, color = Sex))+
  geom_abline(aes(intercept= sum(ncas)/sum(npop)*100, slope =0), linetype=2, color = "black")+
  geom_errorbar(aes(ymin = (est-(est-lower)/1.96) *100, ymax=(est+(upper-est)/1.96)*100), size =0.5, width = 0.3, position = position_dodge(0.5), color = "black")+
  geom_point(aes(fill= Sex), color = "black", size=3, position = position_dodge(0.5), shape =21, stroke=0.5)+
  labs(title = expression("FDC in *a priori* unaffected relatives (*n=536*)"),
       x = "Age")+
  scale_y_continuous(limits = c(0,45), breaks = seq(0,80,5), name = "Prevalence (%)")+
  theme_classic(base_size = 11, base_family = "serif")+
  theme(plot.title =element_markdown(family = "serif", size = 13, color = "black"),
        axis.text = element_text(family = "serif", size = 10, color = "black"),
        axis.title = element_text(family = "serif", size = 10, color = "black"),
        plot.caption = element_text(family = "serif", size = 6),
        legend.position = c(0.15,0.9), legend.title = element_blank(), legend.background = element_blank(),
        panel.grid.major.y = element_line(color = "gray89", size = .3)
        #panel.grid.minor.y = element_line(color = "gray99")
  )+
  scico::scale_fill_scico_d(palette = "batlow", direction = -1)+
  ggpubr::color_palette(palette = chris_col_pal[2:3])+
  geom_curve(aes(x =3.5, xend = 3.7, y=12.5, yend = 17.5), color = "black", arrow = arrow(length = unit(0.02, "npc"), type = "closed"))+
  annotate("text", x=3.3, y = 12.5, label = "Overall\n prevalence", size = 3)+
  coord_cartesian(ylim = c(0,43), xlim = c(.4,4.6), clip = "off", expand = F)
fig2a

ggsave("fig2_a_EHJ-HF.png", units = "cm", width = 12, height = 8, dpi = 2000)
```
**Incidence in men and women**

```{r}
################################################----------------------------------#
#Incidence in men and women relatives
#Evaluated at baseline
#################################################---------------------------------#
spl <-
  as_tibble(survSplit(
    Surv(d3$age_baseline, d3$CM_age + 0.0001, event = d3$CM_c) ~ Sex,
    data = d3,
    cut = c(30, 45, 60),
    episode = "Agegroup"
  )) %>%
  mutate(time = tstop - tstart) %>%
  group_by(Agegroup, Sex) %>%
  summarise(nyear = sum(time), ncas = sum(event)) %>%
  ungroup() %>%
  select(ncas, nyear) %>%
  as.matrix() %>%
  epi.conf(ctype = "inc.rate", method = "byar") %>% 
  bind_cols(
    as_tibble(survSplit(Surv(d3$age_baseline, d3$CM_age + 0.0001, event = d3$CM_c) ~ Sex, data = d3,cut = c(30, 45, 60),episode = "Agegroup")) %>%mutate(time = tstop - tstart) %>% group_by(Agegroup, Sex) %>%  summarise(nyear = sum(time), ncas = sum(event)) %>% ungroup()
  ) %>%
  mutate(
    sur = 1 - est,
    Sex = fct_recode(Sex, women = "F", men = "M"),
    Agegroup = factor(Agegroup, labels = c("<30", "30-44", "45-59", "\u226560"))
  ) %>%
  ggplot(aes(
    x = Agegroup,
    y = est * 1000,
    group = Sex,
    color = Sex
  )) +
  geom_errorbar(aes(ymin = (est - (est - lower) / 1.96) * 1000, ymax = (est + (upper - est) / 1.96) * 1000
    ),
    size = 0.5, width = 0.3,position = position_dodge(0.5),color = "black") +
  geom_point(aes(fill = Sex), color = "black",size = 3, position = position_dodge(0.5), shape = 21, stroke = 0.5
             ) +
  labs(title = "Incidence rate in relatives w/o FDC at baseline (*n=439*)",
       x="Age")+
  scale_y_continuous(
    limits = c(-0.1, 112.5),
    breaks = seq(0, 120, 20),
    name = "Incidence per 1,000 person-years"
  ) +
  theme_classic(base_size = 11, base_family = "serif") +
  theme(
    plot.title = element_markdown(
      family = "serif",
      size = 11,
      color = "black"
    ),
    axis.text = element_markdown(
      family = "serif",
      size = 10,
      color = "black"
    ),
    axis.title = element_markdown(
      family = "serif",
      size = 10,
      color = "black"
    ),
    plot.caption = element_markdown(family = "serif", size = 10),
    legend.position = c(0.15, 0.9),
    legend.title = element_blank(),
    legend.background = element_blank(),
    panel.grid.major.y = element_line(color = "gray89", size = .3)
    #panel.grid.minor.y = element_line(color = "gray99"))+
  ) +
  scale_fill_scico_d(palette = "batlow", direction = -1) +
  scale_color_scico_d(palette = "batlow", direction = -1) +
  
  coord_cartesian(
    xlim = c(0.4, 4.6),
    ylim = c(0, 108),
    clip = "off",
    expand = F
  )



obs_tar<- 
  spl %>%
  select(Agegroup, Sex, ncas) %>%
  pivot_wider(names_from = Agegroup, values_from = ncas) %>%
  bind_rows(
    spl %>% select(Agegroup, Sex, nyear) %>%pivot_wider(names_from = Agegroup, values_from = nyear)
  )
obs_tar
DKstd <- matrix(data = c(2068990,1050499,1544284,1176272), nrow = 1, 
                dimnames = list(c(""), c("<30", "30-44", "45-59", "\u226560")))
ASI <- epi.directadj(obs, tar, DKstd, units = 1, conf.level = 0.95)
```

### Figure 3 
```{r}
#---------------------------------------------#
# Incidence according to objective findings
#---------------------------------------------#
library(ggpubr); library(survminer)
fit1 <- Surv(NoGen$age_baseline, NoGen$CM_age+0.0001, event = NoGen$CM_c)~Abnorm_TTE_or_ECG
spl <- survSplit(fit1, data = NoGen, cut = c(30,45,60), episode = "Agegroup") %>% 
  mutate( time = tstop-tstart,
          ndil = factor(Abnorm_TTE_or_ECG, labels = c("Normal", "Abnormal")),
          Agegroup = factor(Agegroup, labels = c("<30", "30-44", "45-59", "\u226560"))) %>% 
  group_by(Agegroup, ndil) %>% 
  summarise(nyear = sum(time), ncas = sum(event))

tmp <- as.matrix(cbind(spl$ncas, spl$nyear)) %>% 
epi.conf(tmp, ctype = "inc.rate", method = "byar") %>% 
  mutate(ndil = spl$ndil, Agegroup= spl$Agegroup)
chris_col <- scico(6, palette = "batlow")[c(1,4)]
figs <- tmp %>%  ggplot(aes(x=Agegroup, y=est*1000, group = ndil, color = ndil))+
  geom_errorbar(aes(ymin = (est-(est-lower)/1.96)*1000, ymax=(est+(upper-est)/1.96)*1000), size =0.5, width = 0.3, 
                position = position_dodge(0.9), color = "black")+
  geom_point(aes(fill= ndil), color = "black", size=3, position = position_dodge(0.9), shape =21, stroke=0.5)+
  geom_text(aes(label = round(est*1000,0)), size =3, hjust =0.7, vjust = -0.2, color="black", position = position_dodge(0.4))+
  #geom_hline(aes(yintercept = 0), linetype = 2)+
  #labs(title = "Familial dilated cardiomyopathy")+
  xlab("Age")+ labs(fill="ECG & echo")+
  scale_y_continuous(limits = c(-1,120), breaks = seq(0,120,20), name = "Incidence per\n 1,000 person-years")+
  theme_classic(base_size = 12, base_family = "serif")+
  theme(plot.title =element_text(family = "serif", size = 12, color = "black"),
                                 axis.title = element_text(family = "serif", size = 12, colour = "black"),
                                 axis.text = element_text(family = "serif", size = 12, colour = "black"),
                                 axis.ticks = element_line(linetype = "solid"),
                                 axis.line = element_line(),
                                 legend.background = element_blank(),
                                 legend.text = element_text(family = "serif", size = 10),
                                 legend.title = element_text(family = "serif", size = 10),
        legend.position = c(0.15,.8),
                                 plot.background = element_blank(),
                                 panel.background = element_blank(),
                                 panel.grid = element_blank())+ 
  color_palette(palette = chris_col)+  fill_palette(palette = chris_col)+
  coord_cartesian(ylim = c(-1,115), clip = "off", expand = F)
figs
```

```{r}
#---------------------------------------#
#Survival plot
#----------------------------------------#

nogen_plot <- NoGen %>% mutate(FU_diag = if_else(FU_diag>10, 10.01, FU_diag),
                               CM_c = if_else(FU_diag>=10 & CM_c==1, 0,CM_c))
gss <-ggsurvplot(surv_fit(Surv(FU_diag, CM_c)~factor(Abnorm_TTE_or_ECG, levels = c(0,1)), data = nogen_plot),
                 linetype = 1, cenosr.shape = 124, censor.size =3,
                 ylim = c(0,.61),pval = FALSE, pval.coord = c(10,0.82),
                 fun = "event",conf.int = TRUE, conf.int.style = c("ribbon"),conf.int.alpha = 0.2,
                 xlab = "Years of follow-up", #xlim= c(0,13.8),
                 ylab= "Cumulative incidence\n of FDC",
                 break.x.by = 1,
                 break.y.by = .1,
                 legend = c(0.15,0.8), legend.title = "ECG & echo", legend.labs = c("Normal","Abnormal"),
                 palette = c("#001959", "#B28C32" ), 
                 ggtheme = theme(plot.title =element_text(family = "serif", size = 12, color = "black"),
                                 axis.title = element_text(family = "serif", size = 12, colour = "black"),
                                 axis.text = element_text(family = "serif", size = 12, colour = "black"),
                                 axis.ticks = element_line(linetype = "solid"),
                                 axis.line = element_line(),
                                 legend.background = element_blank(),
                                 legend.text = element_text(family = "serif", size = 10),
                                 legend.title = element_text(family = "serif", size = 10),
                                 plot.background = element_blank(),
                                 panel.background = element_blank(),
                                 panel.grid = element_blank()),
                 risk.table=TRUE, risk.table.title = "Numbers at risk", tables.y.text =TRUE, risk.table.pos = "out",
                 tables.height = 0.2, fontsize =3, font.family = "serif",
                 tables.theme = theme_transparent()+theme(text = element_text(family = "serif", size = 8)))
HR_FDC <- exp(coxph(Surv(FU_diag, CM_c)~Abnorm_TTE_or_ECG+age_baseline, data = NoGen)$coef[1])
lwr_FDC <- round(summary(coxph(Surv(FU_diag, CM_c)~Abnorm_TTE_or_ECG+age_baseline, data = NoGen))$conf.int[5],1)
upr_FDC <- round(summary(coxph(Surv(FU_diag, CM_c)~Abnorm_TTE_or_ECG+age_baseline, data = NoGen))$conf.int[7],1)

##########################
COMBINING PLOTS ABOVE
#########################

gss$plot <- gss$plot+
  ggplot2::annotate("text", 
                    x = 1.5, y = 0.3, # x and y coordinates of the text
                    label = paste0( "HR=",(round(HR_FDC,1)),"\n [CI: ",lwr_FDC, " to ", upr_FDC, "]"), 
                    size =4, family = "serif")
grid.draw.ggsurvplot <- function(x){
  survminer:::print.ggsurvplot(x, newpage = FALSE)
}

gss <- grid.draw.ggsurvplot(gss)
grid.draw.ggsurvplot(gss
                     )
ggsave(plot = gss, filename =  "what_1.png", units = "cm", width = 12, height = 10, dpi = 1500)

ggsave(plot=print(gss), filename =  "what2.png", units = "cm", width = 12, height = 8, dpi = 1500)
library(cowplot)
Figs <-plot_grid(figs,gss$plot, ncol = 1, labels = "AUTO")

ggsave("EHJ-HF_figure4_1.png", units = "cm", width = 14, height = 14, dpi = 2000)

```
